%
% concl.tex
%
Our proposed generalized architecture for web-based data visualizaiton meets all outlined objectives as is illustrated by the BeeStream web application.  Our architecture provides sufficient modularity for data sources and datasets, different data scales, data transmission, and charts while remaining general enough to be applicable across various client and server-side web frameworks.  BeeStream not only illustrates the capabilities and practicality of this architecture, but also serves as a basis for easy extension.  BeeStream, and the underlying architecture, met or exceeded our expectations for modularity and adaptability.  \par

Our future work revolves around utilizing our implemneted framework to build visualizations for more complex and varied data and generalizing our solution so that others can easily build applications using our architecture.  A potential extension for BeeStream would be a chart that allows the user to dynamically request datasets, which the driver would then request and serve as defined in our architecture.  Further future work would include building a ``boilerplate'' or starter application that contains the basic components of this architecture.  This starter application should be reproduced in each of the commonly used server and client-side web frameworks in order to accomodate as many developers as possible.  It would also be beneficial to build default \textit{ChartComponents} for each of the major charting library, as library interfaces vary.  Future research can be completed to produce starter code across as many platforms as possible so that others can easily build off of our work.
